---
title: "Module 5: Conditional Statements, Dates, and Tables"
output: learnr::tutorial
runtime: shiny_prerendered
description: "Test your understanding of conditional statements, file naming, and cross-referencing covered in Module 5"
resource_files:
  - ../common/github_usernames.csv
  - ../common/_github_username.Rmd
  - ../common/_participation.Rmd
  - ../common/_submission.Rmd
tutorial:
  id: "md-05-quiz"
---

```{r setup, include=FALSE}
library(learnr)
library(dplyr)
library(palmerpenguins)
library(knitr)
library(gradethis)
library(learnrhash)
library(httr)

# Enable exercise checking
tutorial_options(
  exercise.eval = FALSE,
  exercise.checker = gradethis::grade_learnr
)

knitr::opts_chunk$set(echo = FALSE, warning = FALSE)

# Prepare penguins data
data(penguins, package = "palmerpenguins")
```

## Introduction

This weekly quiz is designed to support your learning in the ds4owd course. It will test your understanding of the concepts covered in Module 5, including conditional statements with `case_when()`, file naming conventions, and cross-referencing tables in Quarto documents. A submission of the quiz for each module is required to complete the course.

## Multiple Choice Questions

### Question 1: File Naming Conventions

```{r q1}
question_checkbox(
  "Which of the following are good practices for naming variables in R? (Select all that apply)",
  answer("Use snake_case for consistency", correct = TRUE),
  answer("Include spaces to make names more readable"),
  answer("Avoid special characters", correct = TRUE),
  answer("Use descriptive but concise names", correct = TRUE),
  answer("Use special characters like () to add information"),
  incorrect = "Good file naming conventions avoid spaces and special characters, and use consistent naming like snake_case.",
  correct = "Correct! Following these naming conventions makes your code more readable and prevents errors.",
  allow_retry = TRUE
)
```

### Question 2: Variable Naming Examples

```{r q2}
question_radio(
  "Which of the following variable names follows best practices for R?",
  answer("Total Solids (g/L)"),
  answer("ts", correct = TRUE),
  answer("Number of users"),
  answer("System (pit latrine / septic tank)"),
  incorrect = "Variable names should avoid spaces, special characters, and be concise. Descriptions belong in a data dictionary.",
  correct = "Correct! 'ts' is a concise variable name without spaces or special characters. The full description ('Total solids in g/L') would go in a data dictionary.",
  allow_retry = TRUE,
  random_answer_order = TRUE
)
```

### Question 3: Table Cross-Referencing

```{r q3}
question_radio(
  "What are the required components to create a cross-referenceable table in a Quarto document?",
  answer("Only a label starting with tbl-"),
  answer("Only a caption with tbl-cap:"),
  answer("Both a label starting with tbl- and a caption", correct = TRUE),
  answer("A label starting with fig- and a caption"),
  incorrect = "Cross-referencing tables requires both a label (starting with tbl-) and a caption in your code chunk options.",
  correct = "Correct! You need both a label (e.g., #| label: tbl-penguins) and a caption (e.g., #| tbl-cap: A few penguins) to make a table cross-referenceable.",
  allow_retry = TRUE,
  random_answer_order = TRUE
)
```

### Question 4: Referencing Tables in Text

```{r q4}
question_radio(
  "If you created a table with the label 'tbl-summary', how would you reference it in your text?",
  answer("@fig-summary"),
  answer("@tbl-summary", correct = TRUE),
  answer("#tbl-summary"),
  answer("[tbl-summary]"),
  incorrect = "To reference a table in text, use @ followed by the label name.",
  correct = "Correct! Use @tbl-summary to create a cross-reference to your table. Quarto will automatically create a numbered reference and hyperlink.",
  allow_retry = TRUE,
  random_answer_order = TRUE
)
```

## Coding Exercises

### Exercise 1: Categorize Penguin Body Mass

Use `mutate()` and `case_when()` to create a new variable `body_mass_category` that categorizes penguins by their body mass:

- Penguins with body mass less than 3500g are "small"
- Penguins with body mass between 3500g and 4500g (inclusive) are "medium"
- Penguins with body mass greater than 4500g are "large"

```{r ex1-setup}
data(penguins, package = "palmerpenguins")
```

```{r ex1, exercise=TRUE, exercise.setup="ex1-setup"}
penguins |>
  mutate(body_mass_category = case_when(
    _______________________________
  ))
```

```{r ex1-solution, include=FALSE}
penguins |>
  mutate(body_mass_category = case_when(
    body_mass_g < 3500 ~ "small",
    body_mass_g >= 3500 & body_mass_g <= 4500 ~ "medium",
    body_mass_g > 4500 ~ "large"
  ))
```

```{r ex1-check}
grade_this_code(
  correct = "Excellent! You've successfully categorized penguins by body mass using case_when().",
  incorrect = "Not quite. {code_feedback()} Remember to check for body_mass_g < 3500, between 3500 and 4500, and > 4500."
)
```

### Exercise 2: Create Island Labels

The penguins dataset contains abbreviated island names. Use `case_when()` to create a new variable `island_full` with full island names:

- "Torgersen" should become "Torgersen Island"
- "Biscoe" should become "Biscoe Island"
- "Dream" should become "Dream Island"

```{r ex2-setup}
data(penguins, package = "palmerpenguins")
```

```{r ex2, exercise=TRUE, exercise.setup="ex2-setup"}
penguins |>
  mutate(island_full = case_when(
    _______________________________
  ))
```

```{r ex2-solution, include=FALSE}
penguins |>
  mutate(island_full = case_when(
    island == "Torgersen" ~ "Torgersen Island",
    island == "Biscoe" ~ "Biscoe Island",
    island == "Dream" ~ "Dream Island"
  ))
```

```{r ex2-check}
grade_this_code(
  correct = "Perfect! You've successfully created full island names using case_when().",
  incorrect = "Not quite. {code_feedback()} You need to match each island name and append ' Island' to it."
)
```

### Exercise 3: Categorize Bill Length by Species

Use `case_when()` to create a new variable `bill_category` that classifies bill length differently for each species:

- For Adelie penguins: bills < 38mm are "short", otherwise "long"
- For Chinstrap penguins: bills < 48mm are "short", otherwise "long"
- For Gentoo penguins: bills < 47mm are "short", otherwise "long"

```{r ex3-setup}
data(penguins, package = "palmerpenguins")
```

```{r ex3, exercise=TRUE, exercise.setup="ex3-setup"}
penguins |>
  mutate(bill_category = case_when(
    _______________________________
  ))
```

```{r ex3-solution, include=FALSE}
penguins |>
  mutate(bill_category = case_when(
    species == "Adelie" & bill_length_mm < 38 ~ "short",
    species == "Adelie" & bill_length_mm >= 38 ~ "long",
    species == "Chinstrap" & bill_length_mm < 48 ~ "short",
    species == "Chinstrap" & bill_length_mm >= 48 ~ "long",
    species == "Gentoo" & bill_length_mm < 47 ~ "short",
    species == "Gentoo" & bill_length_mm >= 47 ~ "long"
  ))
```

```{r ex3-check}
grade_this_code(
  correct = "Excellent work! You've successfully applied different conditions for each species.",
  incorrect = "Not quite. {code_feedback()} Remember to check both species AND bill_length_mm for each condition."
)
```

### Exercise 4: Fix Data Entry Error

A data entry error occurred for penguin observations from Dream island in 2007. All Adelie penguins from Dream island in 2007 had their flipper length recorded incorrectly - it's 10mm too short. Use `case_when()` to add 10mm to the flipper length only for these specific observations, keeping all other values unchanged.

Hint: Use `.default = flipper_length_mm` to keep unchanged values.

```{r ex4-setup}
data(penguins, package = "palmerpenguins")
```

```{r ex4, exercise=TRUE, exercise.setup="ex4-setup"}
penguins |>
  mutate(flipper_length_mm = case_when(
    _______________________________
  ))
```

```{r ex4-solution, include=FALSE}
penguins |>
  mutate(flipper_length_mm = case_when(
    species == "Adelie" & island == "Dream" & year == 2007 ~ flipper_length_mm + 10,
    .default = flipper_length_mm
  ))
```

```{r ex4-check}
grade_this_code(
  correct = "Perfect! You've successfully corrected the data entry error using multiple conditions.",
  incorrect = "Not quite. {code_feedback()} You need to check for species == 'Adelie', island == 'Dream', AND year == 2007, then add 10 to flipper_length_mm. Don't forget .default to keep other values unchanged."
)
```

## Details

**Important**: Please enter your GitHub username below and click "Submit Quiz" at the end to record your completion.

```{r child="_github_username.Rmd"}
```

```{r child="_participation.Rmd"}
```

```{r child="_submission.Rmd"}
```