---
title: "Module 1: Getting Started"
output: learnr::tutorial
runtime: shiny_prerendered
description: "Test your understanding of the concepts covered in Module 1" 
resource_files:
  - github_usernames.csv
  - _github_username.Rmd
  - _participation.Rmd
  - _submission.Rmd
tutorial:
  id: "md-01-quiz"
---

```{r setup, include=FALSE}
library(learnr)
library(dplyr)
library(gapminder)
library(knitr)
library(gradethis)
library(learnrhash)
library(httr)

# Enable exercise checking
tutorial_options(
  exercise.eval = FALSE,
  exercise.checker = gradethis::grade_learnr
)

knitr::opts_chunk$set(echo = FALSE)

```

## Introduction

This weekly quiz is designed to support your learning in the ds4owd course. It will test your understanding of the concepts covered in Module 1, including RStudio IDE, GitHub Personal Access Tokens (PATs), and basic R syntax. A submission of the quiz for each module is required to complete the course. 

## Multiple Choice Questions

### Question 1: GitHub PAT Storage

```{r q1}
question_radio(
  "Where should you store your GitHub Personal Access Token (PAT) for secure and convenient use when working in cloud environments like Posit Cloud?",
  answer("In a text file on your desktop for easy access"),
  answer("In your R script as a comment for reference"),
  answer("In a secure password manager or credentials manager", correct = TRUE),
  answer("In your browser's bookmarks for quick retrieval"),
  incorrect = "GitHub PATs should be stored securely in a password manager or credentials manager to prevent unauthorized access. Since we'll be working in Posit Cloud, system credential managers aren't available, so external password managers are recommended.",
  correct = "Correct! Storing PATs in a secure password manager or credentials manager provides security while allowing you to access them when needed in cloud environments where system credential stores aren't available.",
  allow_retry = TRUE,
  random_answer_order = TRUE
)
```

### Question 2: RStudio IDE Panes

```{r q2}
question_checkbox(
  "Which four panes make up the RStudio IDE? (Select all four)",
  answer("Source Editor", correct = TRUE),
  answer("Console", correct = TRUE),
  answer("Environment/History/Git/Connections", correct = TRUE),
  answer("Files/Plots/Packages/Help/Viewer", correct = TRUE),
  answer("Terminal"),
  answer("Debug"),
  answer("Version Control"),
  incorrect = "The RStudio IDE has four main panes. Try again and select exactly four options.",
  correct = "Excellent! The four panes are: Source Editor (top-left), Console (bottom-left), Environment/History/Connections (top-right), and Files/Plots/Packages/Help/Viewer (bottom-right).",
  allow_retry = TRUE
)
```

### Question 3: R Assignment Operator

```{r q5}
question_radio(
  "Which is the assignment operator in R?",
  answer("="),
  answer("<-", correct = TRUE),
  answer("<<-"),
  answer("->"),
  incorrect = "The tidyverse style guide recommends using a specific assignment operator for clarity and consistency.",
  correct = "Correct! The arrow operator '<-' is preferred in R for assignment, as it clearly shows the direction of assignment and avoids confusion with function arguments.",
  allow_retry = TRUE,
  random_answer_order = TRUE
)
```

## Coding Exercises

### Exercise 1: Modifying gapminder data

The code below filters the gapminder dataset for the year 2007 and calculates the mean life expectancy by continent. Modify the code to filter for the year 1952 instead.

```{r exercise1, exercise=TRUE}
gapminder |> 
  filter(year == 2007) |> 
  group_by(continent) |>
  summarise(
    mean_life_exp = mean(lifeExp)
  )
```

```{r exercise1-solution}
gapminder |> 
  filter(year == 1952) |> 
  group_by(continent) |>
  summarise(
    mean_life_exp = mean(lifeExp)
  )
```

```{r exercise1-check}
grade_this_code()
```

## Details

**Important**: Please enter your GitHub username below and click "Submit Quiz" at the end to record your completion.

```{r github-username, child='_github_username.Rmd'}
```

```{r participation, child='_participation.Rmd'}
```

```{r submission-section, child='_submission.Rmd'}
```
  
You've completed the Module 1 quiz! 
